---
title: "OCD_CTD_TADA.qmd"
last updated: 20251006
format: html
editor: visual
---

# Set up workspace

```{r}
library(tidyverse)
library(readxl)

#import functions
source("TADAfunctions.R")

#import TADA input tables (per cohort, tables include priors, and cohort-level dn/rare variant counts)
tada_input_CTD <- read_excel("OCD_CTD_TADA_input.xlsx",
    sheet = "tada_input_CTD")
tada_input_OCD <- read_excel("OCD_CTD_TADA_input.xlsx",
    sheet = "tada_input_OCD")
tada_input_OCDCTD <- read_excel("OCD_CTD_TADA_input.xlsx",
    sheet = "tada_input_OCDCTD")

#import gene constraint and mutability metrics
mutTable <- read_excel("OCD_CTD_TADA_input.xlsx",
    sheet = "mutTable")
```

# CTD

```{r}
#dnCTD TADA
n_case <- 1550
n_con  <- 1734

#CTD pi
pi = 0.05848704

#TADA, dnCTD, poisson burden
tada_dnCTD_pois <- tada_input_CTD %>%
  mutate(
    BF_dnLGD = BF_DN_SNV(
      count_case = CTD_dnLGD,
      count_con  = SSC_dnLGD,
      n_case     = n_case,
      n_con      = n_con,
      mut        = gnomad.dnLGD.mu,
      gamma.dn   = gamma_pois_dnLGD,
      beta.dn    = 0.2
    ),
    BF_dnMis3 = BF_DN_SNV(
      count_case = CTD_dnMis3,
      count_con  = SSC_dnMis3,
      n_case     = n_case,
      n_con      = n_con,
      mut        = gnomad.dnMis3.mu,
      gamma.dn   = gamma_pois_dnMis3,
      beta.dn    = 0.2
    )
  ) %>% 
  mutate(BF = BF_dnLGD * BF_dnMis3) %>% 
  mutate(
    qval=Bayesian.FDR(BF, pi0 = 1-pi)) %>% 
  mutate(p = q2p(qval, gene)) %>% 
  arrange(qval)

tada_dnCTD_pois %>%
  filter(qval < 0.3) %>%
  mutate(
    qval = signif(qval, 2),
    p = signif(p, 2),
    LOEUF = signif(LOEUF, 2),
    misZ = signif(misZ, 2)
  ) %>%
  dplyr::select(gene, qval, p, pLI, LOEUF, misZ, CTD_dnLGD, CTD_dnMis3, SSC_dnLGD, SSC_dnMis3)
```

# OCD

```{r}
#dnOCD TADA
n_case <- 1493
n_con  <- 1734
n_sing_case = 1546
n_sing_ctrl = 2380

#OCD pi
pi = 0.0294039

tada_dnccOCD_pois <- tada_input_OCD %>%
  mutate(
    BF_dnLGD = BF_DN_SNV(
      count_case = OCD_dnLGD,
      count_con  = SSC_dnLGD,
      n_case     = n_case,
      n_con      = n_con,
      mut        = gnomad.dnLGD.mu,
      gamma.dn   = gamma_pois_dnLGD,
      beta.dn    = 0.2
    ),
    BF_dnMis3 = BF_DN_SNV(
      count_case = OCD_dnMis3,
      count_con  = SSC_dnMis3,
      n_case     = n_case,
      n_con      = n_con,
      mut        = gnomad.dnMis3.mu,
      gamma.dn   = gamma_pois_dnMis3,
      beta.dn    = 0.2
    ),
    BF_rareLGD = BF_CC_SNV(
      count_case = OCD_rareLGD, 
      count_con = SSC_rareLGD,
      n_case = n_sing_case,
      n_con = n_sing_ctrl,
      mut = gnomad.rareLGD.mu,
      gamma.cc = gamma_nb_rareLGD,
      nu=5000)) %>% 
  mutate(BF = BF_dnLGD * BF_dnMis3 * BF_rareLGD) %>% 
  mutate(
    qval=Bayesian.FDR(BF, pi0 = 1-pi)) %>% 
  mutate(p = q2p(qval, gene)) %>% 
  arrange(qval)

tada_dnccOCD_pois %>% filter(qval<0.1) %>% dplyr::select(gene, pLI, LOEUF, misZ, SSC_dnMis3:SSC_rareLGD, qval, p)

tada_dnccOCD_pois %>% 
  filter(qval < 0.3) %>%
  mutate(
    qval = signif(qval, 2),
    p = signif(p, 2),
    LOEUF = signif(LOEUF, 2),
    misZ = signif(misZ, 2)
  ) %>%
  dplyr::select(gene, qval, p, pLI, LOEUF, misZ, OCD_dnLGD, OCD_dnMis3, OCD_rareLGD, SSC_dnLGD, SSC_dnMis3, SSC_rareLGD)
```

# OCDCTD

```{r}
#dnccOCDCTD TADA
n_case <- 2418
n_con  <- 1734
n_sing_case = 1546
n_sing_ctrl = 2380

#OCDCTD pi
pi = 0.04063085


tada_dnccOCDCTD_pois <- tada_input_OCDCTD %>%
  mutate(
    BF_dnLGD = BF_DN_SNV(
      count_case = OCDCTD_dnLGD,
      count_con  = SSC_dnLGD,
      n_case     = n_case,
      n_con      = n_con,
      mut        = gnomad.dnLGD.mu,
      gamma.dn   = gamma_pois_dnLGD,
      beta.dn    = 0.2
    ),
    BF_dnMis3 = BF_DN_SNV(
      count_case = OCDCTD_dnMis3,
      count_con  = SSC_dnMis3,
      n_case     = n_case,
      n_con      = n_con,
      mut        = gnomad.dnMis3.mu,
      gamma.dn   = gamma_pois_dnMis3,
      beta.dn    = 0.2
    ),
    BF_rareLGD = BF_CC_SNV(
      count_case = OCD_rareLGD, 
      count_con = SSC_rareLGD,
      n_case = n_sing_case,
      n_con = n_sing_ctrl,
      mut = gnomad.rareLGD.mu,
      gamma.cc = gamma_nb_rareLGD,
      nu=5000))


tada_dnccOCDCTD_pois = tada_dnccOCDCTD_pois %>% 
  mutate(BF = BF_dnLGD * BF_dnMis3 * BF_rareLGD) %>% 
  mutate(
    qval=Bayesian.FDR(BF, pi0 = 1-pi)) %>% 
  mutate(p = q2p(qval, gene)) %>% 
  arrange(qval)

#look at hcOCDCTD
tada_dnccOCDCTD_pois %>% 
  filter(qval < 0.1) %>%
  mutate(
    qval = signif(qval, 2),
    p = signif(p, 2),
    LOEUF = signif(LOEUF, 2),
    misZ = signif(misZ, 2)
  ) %>%
  dplyr::select(gene, qval, p, pLI, LOEUF, misZ, OCDCTD_dnLGD, OCDCTD_dnMis3, OCD_rareLGD, SSC_dnLGD, SSC_dnMis3, SSC_rareLGD)
```

# Compile TADA results

Output: tada_results dataframe

```{r}
temp <- list(
  dnCTD      = tada_dnCTD_pois,
  dnccOCD   = tada_dnccOCD_pois,
  dnccOCDCTD = tada_dnccOCDCTD_pois
)

keep_pattern <- paste(c("qval", "SSC", "OCD",  "CTD", "OCDCTD"), collapse = "|")

# Rename qval column in each dataframe, then select relevant columns
temp2 <- imap(temp, ~ 
  .x %>%
    ## Rename 'qval' column in each dataframe to e.g., 'q_dnOCD'
    dplyr::rename(!!paste0("q_", .y) := qval) %>%
    dplyr::select(gene, matches(keep_pattern))
)

# Perform full join across all renamed dataframes
tada_results <- purrr::reduce(temp2, full_join, by = "gene") %>%
  #deduplicate
  dplyr::select(all_of(names(.)[!duplicated(gsub("\\.x+|\\.y+", "", names(.)))])) %>%
  # Clean column names by removing .x, .y, .x.x, etc.
  setNames(gsub("\\.x+|\\.y+", "", names(.))) %>%
  #Reorders columns as:gene, q_, other columns in alphabetical order, ssc columns last
  {
    all_cols <- names(.)
    q_cols <- grep("^q_", all_cols, value = TRUE)
    ssc_cols <- grep("^SSC_", all_cols, value = TRUE)
    other_cols <- setdiff(all_cols, c("gene", q_cols, ssc_cols))
    dplyr::select(., gene, all_of(q_cols), all_of(sort(other_cols)), all_of(sort(ssc_cols)))
  } %>%
  mutate(gamma_class = "pois") %>% dplyr::relocate(gamma_class)  %>%
  # Join with mutTable
  full_join(
    mutTable %>% dplyr::select(gene, transcript, LOEUF, LOEUF_decile, pLI, misZ, gnomad.dnLGD.mu:gnomad.rare_mis3.mu),
    by = "gene"
  ) %>%
  dplyr::select(-gamma_class, -gnomad.rareMisA.mu, -gnomad.rareMisB.mu, -gnomad.dnMisB.mu, -gnomad.rare_mis3.mu) %>%
  dplyr::select(gene, transcript, LOEUF, LOEUF_decile, pLI, 
                gnomad.dnLGD.mu, gnomad.dnMis3.mu, gnomad.rareLGD.mu,
                q_dnccOCD, q_dnCTD, q_dnccOCDCTD,
                CTD_dnLGD:SSC_rareLGD) %>%
  arrange(q_dnccOCDCTD)

#save results as excel file
#write.xlsx(tada_results, file = "output_path/OCDCTD_tada_results.xlsx")
```

## How many significant hc/p genes?

Tally the number of significant hc/p genes

-   hc: FDR \< 0.1

-   p: FDR 0.1-0.3 and at least 2 case dnDam

```{r}
nGenes = tada_results %>%
  transmute(
    gene,
    dnCTD = q_dnCTD,
    dnccOCD = q_dnccOCD,
    dnccOCDCTD = q_dnccOCDCTD,
    CTD_dnDamaging = CTD_dnLGD + CTD_dnMis3,
    OCD_dnDamaging = OCD_dnLGD + OCD_dnMis3,
    OCDCTD_dnDamaging = OCDCTD_dnLGD + OCDCTD_dnMis3,
  ) %>%
  pivot_longer(
    cols = starts_with("dn"),
    names_to = "model",
    values_to = "qval"
  ) %>%
  mutate(
    dnDamaging = case_when(
      model == "dnCTD" ~ CTD_dnDamaging,
      model == "dnccOCD" ~ OCD_dnDamaging,
      model == "dnccOCDCTD" ~ OCDCTD_dnDamaging
      )
  ) %>%
  group_by(model) %>%
  summarise(
    hc = sum(qval < 0.1, na.rm = TRUE),  # no dnDamaging filter
    p = sum(qval >= 0.1 & qval < 0.3 & dnDamaging >= 2, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(model = factor(
    model,
    levels = c( "dnCTD","dnccOCD", "dnccOCDCTD")
  )) %>%
  arrange(model)
nGenes

    #   model         hc     p
    #   <fct>      <int> <int>
    # 1 dnCTD         10     6
    # 2 dnccOCD       12    16
    # 3 dnccOCDCTD    34    20
```
